---
author: Gene Leynes
date: '2021-08-30'
title: Example - Changes in 2020 Tracts from 2010
output: html_document
editor_options: 
  chunk_output_type: console
---

These maps are simple outlines of the tracts for 2010 and 2020. 


```{r, message=FALSE}
## Libraries

library(geneorama)

# library(shiny)
library(leaflet)
library(RColorBrewer)
library(rgdal)
library(rgeos)
library(sp)
library(data.table)
library(colorspace)

## API Key for stadia
leaflet_stadia_url <- "https://tiles.stadiamaps.com/tiles/alidade_smooth_dark/{z}/{x}/{y}{r}.png?apikey={'%s'}" %>% sprintf(., yaml::read_yaml("config/stadia.yaml"))
leaflet_stadia_attribution <- '&copy; <a href="https://stadiamaps.com/">Stadia Maps</a>, &copy; <a href="https://openmaptiles.org/">OpenMapTiles</a> &copy; <a href="http://openstreetmap.org">OpenStreetMap</a> contributors'
```

## Shape files

The steps:

 - Read in shape files for 2010 and 2020
 - Examine the data portion
 - Geocode tracts to community area (just to determine if they're in the City of Chicago)
 - Use the CA map to generate an outline of the city as well, just for mapping (not always used)


```{r}

## Shape files
shp_tracts_2020_10 <- readOGR("data/tl_2020_17031_tract10")
shp_tracts_2020_20 <- readOGR("data/tl_2020_17031_tract20")

## Look at the data format 
str(shp_tracts_2020_10@data)
str(shp_tracts_2020_20@data)

## Bring in CA data and geocode tracts
data("chi_community_areas")
shp_tracts_2020_10$community <- geocode_to_map(lat = shp_tracts_2020_10$INTPTLAT10, 
                                               lon = shp_tracts_2020_10$INTPTLON10,
                                               map = chi_community_areas,
                                               "community")
shp_tracts_2020_20$community <- geocode_to_map(lat = shp_tracts_2020_20$INTPTLAT20, 
                                               lon = shp_tracts_2020_20$INTPTLON20,
                                               map = chi_community_areas,
                                               "community")

## Determine if in Chicago
shp_tracts_2020_10 <- shp_tracts_2020_10[!is.na(shp_tracts_2020_10@data$community), ]
shp_tracts_2020_20 <- shp_tracts_2020_20[!is.na(shp_tracts_2020_20@data$community), ]

## Generate city outline
chi_city_outline <- rgeos::gUnaryUnion(as(chi_community_areas, "SpatialPolygons"))
```


## Pick a subset of the city as example

I'm really just interested in understanding generally what happened by looking at a section of the city visually. This step is just to look at the Community Areas in order to pick a subset of tracts that are just big enough to be meaningful. 


```{r}
## SAMPLE COMMUNITY AREAS TO PLOT FASTER
# samp_ca_num <- 1:77  ## To show all CA's, use this line instead
samp_ca_num <- c(8,24,28,31,32,33,34,35,59,60)
samp_ca_names <- chi_community_areas$community[chi_community_areas$area_numbe %in% samp_ca_num]

## To show all CA's, use this line instead
# samp_ca_num <- 1:77  
# samp_ca_names <- chi_community_areas$community[chi_community_areas$area_numbe %in% samp_ca_num]


## Get colors for each community area using 4 color algorithm
pal <- colorspace::qualitative_hcl(n = 4, h = c(26, -264), c = 70, l = 70)
chi_community_areas$colors <- pal[MapColoring::getColoring(chi_community_areas)]



leaflet(chi_community_areas[chi_community_areas$area_numbe %in% samp_ca_num, ]) %>%
    addTiles(urlTemplate = leaflet_stadia_url, 
             attribution = leaflet_stadia_attribution) %>% 
    addPolygons(fillColor = ~colors, fillOpacity = 0.5,
                weight = 0.5, label = ~paste(area_numbe, ":", community) ) %>% 
    addLabelOnlyMarkers(~lon_centroid, 
                        ~lat_centroid, label = ~paste(area_numbe, ":", community),
                        labelOptions = labelOptions(noHide = TRUE,
                                                    direction = "center",
                                                    offset = c(0, 0), opacity = 1, 
                                                    textsize = "10px", textOnly = TRUE, 
                                                    style = list("font-style" = "bold")))

```

## Compare 2020 to 2010

Using label offsets helps show the 2020 tract labels next the 2010 labels.

This map shows:

 - There were a lot of tracts with minor, but insignificatnt edits. 
 - The tract numbers are consistent, at least a lot of the time
 

```{r}

shp_tracts_2020_10_subset <- shp_tracts_2020_10[shp_tracts_2020_10$community %in% samp_ca_names, ]
shp_tracts_2020_20_subset <- shp_tracts_2020_20[shp_tracts_2020_20$community %in% samp_ca_names, ]

## Zoomed in view of subset
leaflet() %>%
    addTiles(urlTemplate = leaflet_stadia_url, 
             attribution = leaflet_stadia_attribution) %>% 
    addPolygons(data=shp_tracts_2020_10_subset,
                color="blue", fill = FALSE) %>%
    addLabelOnlyMarkers(data = shp_tracts_2020_10_subset,
                        lat = ~INTPTLAT10 %>% as.numeric, 
                        lng = ~INTPTLON10 %>% as.numeric,
                        label = ~TRACTCE10,
                        labelOptions = labelOptions(noHide = TRUE,
                                                    direction = "center",
                                                    offset = c(0, 5), opacity = 1, 
                                                    textsize = "10px", textOnly = TRUE, 
                                                    style = list("color" = "white",
                                                                 "font-style" = "bold")))%>%
    addPolygons(data=shp_tracts_2020_20_subset,
                color="red", fill = FALSE) %>%
    addLabelOnlyMarkers(data = shp_tracts_2020_20_subset,
                        lat = ~INTPTLAT20 %>% as.numeric, 
                        lng = ~INTPTLON20 %>% as.numeric,
                        label = ~TRACTCE20,
                        labelOptions = labelOptions(noHide = TRUE,
                                                    direction = "center",
                                                    offset = c(0, -5), opacity = 1, 
                                                    textsize = "10px", textOnly = TRUE, 
                                                    style = list("color" = "white",
                                                                 "font-style" = "bold")))
```