---
author: Gene Leynes
date: '2021-08-30'
title: Diversity Index by Tract
output: html_document
editor_options: 
  chunk_output_type: console
---


The diversity index for Chicago


```{r, message=FALSE, results='hide'}

## Libraries
library(geneorama)
library(shiny)
library(leaflet)
library(RColorBrewer)
library(rgdal)
library(rgeos)
library(sp)
library(data.table)
library(colorspace)

## API Key for stadia
leaflet_stadia_url <- "https://tiles.stadiamaps.com/tiles/alidade_smooth_dark/{z}/{x}/{y}{r}.png?apikey={'%s'}" %>% sprintf(., yaml::read_yaml("config/stadia.yaml"))
leaflet_stadia_attribution <- '&copy; <a href="https://stadiamaps.com/">Stadia Maps</a>, &copy; <a href="https://openmaptiles.org/">OpenMapTiles</a> &copy; <a href="http://openstreetmap.org">OpenStreetMap</a> contributors'

sourceDir("R/functions/")
```


## Read in demographic data

Read in only the fields from `data/il2020.pl_COMBINED_TRACT.csv` that are needed for calculating the diversity index, and use the function to calculate the field. 

The diversity index indicates how likely you are to pick people of two different races when you pick two people from a population. It's calculated by calculating the propbability that you don't pick two people of the same race. 

To calculate race in this context the variables that break out population by Hispanic / Non-Hispanic are used, and for simplicity of calculation the (relatively small) multiracial population is all treated as a single race. This simplifying assumption has little effect on the result, and makes the index much easier to compute. Plus, since it's consistently computed, it's reasonable to use as a comparison between regions.

Relevant fields for Diversity Index:
    
 - `P0020001` - Total pop
 - `P0020002` - Hispanic or Latino
 - `P0020005` - Not Hispanic or Latino: White alone
 - `P0020006` - Not Hispanic or Latino:  Black or African American alone 
 - `P0020007` - Not Hispanic or Latino:  American Indian and Alaska Native alone
 - `P0020008` - Not Hispanic or Latino:  Asian alone
 - `P0020009` - Not Hispanic or Latino:  Native Hawaiian and Other Pacific Islander alone
 - `P0020010` - Not Hispanic or Latino:  Some Other Race alone
 - `P0020011` - Not Hispanic or Latino:  Population of two or more races
    
Other header fields: `GEOID`, `GEOCODE`, `STATE`, `COUNTY`, `SUMLEV`, `TRACT`, `BLKGRP`, 
               `BLOCK`, `CSA`, `BASENAME`, `POP100`, `HU100`, 
               `INTPTLAT`, `INTPTLON` 


Formula for Diversity Index: `1 - ((H/TOT)^2 + (W/TOT)^2 + (B/TOT)^2 + (AIAN/TOT)^2 + 
                          (ASIAN/TOT)^2 + (NHPI/TOT)^2 + (SOR/TOT)^2 + 
                          (MULTI/TOT)^2)`

The logic for these computations is encapsulated in the R functions in `R/functions`. 


```{r}
pl_2020_tract <- "data/il2020.pl_COMBINED_TRACT.csv" %>% 
    fread_census %>%
    di_race_var_table
pl_2020_tract$DI <- diversity_index(pl_2020_tract)
pl_2020_tract
```

A histogram of the index:

```{r}
hist(pl_2020_tract$DI)
```



```{r, results='hide'}
## Import shape file
shp_tracts_2020 <- readOGR("data/tl_2020_17031_tract20")

## Clean the year specific column names (i.e. ALAND20 becomes ALAND)
shp_tracts_2020@data %>% setnames(., gsub("[1-2]0", "", colnames(.)))


## Some work to look for a common key

## View data structure
str(shp_tracts_2020@data)
str(pl_2020_tract)

NAsummary(shp_tracts_2020@data)
NAsummary(pl_2020_tract)
# data.table(shp_tracts_2020@data$GEOID %>% sort,pl_2020_tract$GEOCODE %>% sort)[,list(.N),V1==V2]

## Bring in CA data and geocode tracts
data("chi_community_areas")

## Generate city outline
chi_city_outline <- rgeos::gUnaryUnion(as(chi_community_areas, "SpatialPolygons"))


## Geocode to CA
shp_tracts_2020$community <- geocode_to_map(lat = shp_tracts_2020$INTPTLAT, 
                                            lon = shp_tracts_2020$INTPTLON,
                                            map = chi_community_areas,
                                            "community")
## Determine if geogrpahies are in Chicago
shp_tracts_2020$chicago <- !is.na(shp_tracts_2020@data$community)

##==============================================================================
## COMMUNITY AREA DATA / GEOCODING
##==============================================================================
setnames(shp_tracts_2020@data,"GEOID", "GEO_ID")
shp_tracts_2020@data <- merge(shp_tracts_2020@data,
                              pl_2020_tract[, list(GEO_ID, DI)], 
                              "GEO_ID", sort = FALSE)
```


```{r}
pal <- viridisLite::viridis(256, option = "D") %>% 
    colorNumeric(palette = ., 
                 domain = shp_tracts_2020$DI,
                 na.color = "black", alpha = F, reverse = TRUE)

leaflet(shp_tracts_2020) %>%
    addTiles(urlTemplate = leaflet_stadia_url,
             attribution = leaflet_stadia_attribution) %>%
    addPolygons(fillColor = ~pal(DI),
                fillOpacity = 0.5, 
                stroke=FALSE, weight = 0.5,
                label=~paste(GEO_ID, DI)) %>%
    addLegend(pal = pal,
              values = ~DI,
              title = "Diversity Index",
              position = "bottomright")
```


```{r}

library(viridisLite)

# shp <- dat[Hard_break_event>0,
#            list(intersectionId, intersectionName,
#                 latitude, 
#                 longitude, 
#                 hard_break_event = Hard_break_event)]
# shp[ , cols := colorNumeric("Greens", NULL, n = 4)(log(hard_break_event))]
# shp[ , popups := paste(intersectionId, intersectionName)]
# leaflet() %>%
#     addProviderTiles("CartoDB.Positron") %>%
#     addCircleMarkers(lng = ~ longitude, lat = ~ latitude, 
#                      data = shp, popup = ~popups,
#                      radius = 3, color = ~cols, stroke = FALSE, 
#                      fillOpacity = 0.5) 
#     # addLegend(pal = ctpal, values = log(preds), ## NOTE: USING LOG VALUES
#     #           position = "bottomright", title = "Vector Index Values")

viridisLite::viridis(256, option = "D") %>% 
    colorNumeric(palette = ., 
                 domain = "",
                 na.color = "black", alpha = F, reverse = TRUE)

palette_function <- colorNumeric(palette = "Greens", NULL, n = 4)
palette_function <- colorNumeric(palette = "inferno", NULL, n = 4, reverse = T)
shp <- dat[Hard_break_event>0,
           list(intersectionId, intersectionName,
                latitude, 
                longitude, 
                Hard_break_event,
                cols = palette_function((log(Hard_break_event))),
                popups = paste(intersectionId, intersectionName))]
leaflet(shp) %>%
    addProviderTiles("CartoDB.Positron") %>%
    addCircleMarkers(lng = ~ longitude, lat = ~ latitude, popup = ~popups,
                     radius = 4, color = ~cols, stroke = FALSE, 
                     label = ~popups,
                     fillOpacity = 0.25) %>% 
    addLegend(pal = palette_function,
              values = ~Hard_break_event %>% log,
              position = "bottomright",
              title = "Hard Brake Events")



palette_function <- colorNumeric(palette = "inferno", NULL, n = 4, reverse = T)
leaflet(shp) %>%
    addProviderTiles("CartoDB.Positron") %>%
    addCircleMarkers(lng = ~ longitude, lat = ~ latitude, popup = ~popups,
                     radius = 8, color = ~cols, stroke = FALSE, 
                     fillOpacity = 0.25) %>% 
    addLegend(pal = palette_function,
              values = ~Hard_break_event %>% log,
              position = "bottomright",
              title = "Hard Brake Events") %>% 
    fitBounds(-87.5869560241699, 41.8921189164765, -87.6624870300293, 41.8665567412712)
```

